---
title: "fitnessdevice"
author: "Patricia"
date: "October 9, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Fitness Device Analysis

First, load input data

```{r readdata}
setwd("C:/Users/trish/Desktop/coursera/rprogram/reproducible research")
fitness<-read.csv(file="activity.csv", header=TRUE, sep=",")

```
## plot histogram of daily steps

```{r plothist}
dailytotal<-aggregate(x=fitness$steps, by=list(fitness$date), FUN=sum)
names(dailytotal)<-c('date', 'steps')
hist(dailytotal$steps)

```

## Calculate mean and median daily statistics

```{r calcstats}
dailymean<-mean(dailytotal$steps, na.rm=TRUE)
dailymedian<-median(dailytotal$steps, na.rm=TRUE)

```
Average daily steps is `r  dailymean`.  

Median daily steps is `r dailymedian`.  


## Including Plots

This is a time series of the daily steps:  

```{r timeseries, echo=FALSE}
plot(dailytotal,main="Time Series of Daily Steps", rm.na=TRUE)
lines(dailytotal)
```

## STEP 5 
```{r maxinterval, echo=FALSE}
  library(dplyr)
    getmode <- function(v) {
        uniqv <- unique(v)
        uniqv[which.max(tabulate(match(v, uniqv)))]
                           }
  fitness1<-fitness
  maxinterval<-fitness1%>% group_by(date)%>%filter(steps==max(steps))
  mdint<-getmode(maxinterval$interval)
```

The 5 minute interval with the most steps on average is `r mdint`.  

## Step 6
```{r treatna, echo=FALSE}
        impute <- function(x, fun) {
        missing <- is.na(x)
        replace(x, missing, fun(x[!missing]))
                              }  
         fitness2<-fitness
        fitness2$interval<-as.factor(fitness2$interval)
        meanstepstb<-fitness2 %>% group_by(interval)%>%summarise(meanint=mean(as.numeric(steps),na.rm=TRUE) )
      
   fitness_nona<-fitness2 %>%group_by(interval) %>% mutate(steps = impute(steps, mean) )
   cleandailymean<-fitness_nona%>%group_by(date)%>%summarise(dailystep=sum(steps))
```
 I replaced na values with mean value for the relevant interval.
 
## Step 7 
Plot histogram of cleand data
```{r plotclean, echo=FALSE}
        hist(cleandailymean$dailystep)
```
 
## Step 8 Panel plot of average per interval

```{r panelplot, echo=FALSE}
    fitness4<-fitness_nona
    fitness4$weekend = chron::is.weekend(fitness4$date)
    s4<- fitness4%>% group_by(interval, weekend)%>%summarise(intervalmean=mean(steps))
    s4<-s4%>% mutate(wl=if_else(weekend, "Weekend", "Weekday")) 
    xyplot(s4$intervalmean~as.numeric(s4$interval)|s4$wl, main="Ave steps per interval for weekdays and weekends", 
            ylab="Average steps in interval", xlab="interval")
```

